### 2023-03-21 chapter0/chapter0.tex
Added citations for studies of activation domains

Both transcription and translation are tightly regulated processes, but because cells employ numerous interlocking mechanisms to control access to DNA and the creation of mRNA, transcription is often the major regulatory checkpoint in gene expression. A key step in this process is the recruitment of the protein complex which transcribes DNA into mRNA to the beginning of a gene's sequence, \textit{i.e.} its transcription start site (TSS). The placement of this complex, called RNA polymerase, is dictated by regulatory sequences encoded in the DNA, and while some regulatory sequences are ubiquitous, like those that mark a gene's TSS, many are gene-specific. Thus, the unique collection of regulatory sequences associated with a gene determines many aspects of its expression. RNA polymerase does not directly contact these gene-specific regulatory elements, however. Instead, proteins called transcription factors directly bind these sequences and interact with other components of the transcriptional machinery which in turn recruit RNA polymerase to the TSS. Accordingly, a typical transcription factor has two parts: a DNA-binding domain and an activation domain. DNA-binding domains are usually structured because they make stable contacts with specific regulatory sequences. In contrast, activations domains interact with the transcriptional machinery and are highly enriched in IDRs. Decades of research have identified many of the proteins involved in these interactions as well as some common properties of their sequences. Despite this progress, however, the precise mechanisms by which transcription factors locate their DNA targets and recruit the transcriptional machinery to initiate expression of specific genes remains an unsolved problem in molecular biology. As disordered regions play a key role in this process, studies which broadly investigate the relationship between the sequence, structure, and function of IDRs therefore have significant implications for our understanding of gene regulation.

>>>>

Both transcription and translation are tightly regulated processes, but because cells employ numerous interlocking mechanisms to control access to DNA and the creation of mRNA, transcription is often the major regulatory checkpoint in gene expression. A key step in this process is the recruitment of the protein complex which transcribes DNA into mRNA to the beginning of a gene's sequence, \textit{i.e.} its transcription start site (TSS). The placement of this complex, called RNA polymerase, is dictated by regulatory sequences encoded in the DNA, and while some regulatory sequences are ubiquitous, like those that mark a gene's TSS, many are gene-specific. Thus, the unique collection of regulatory sequences associated with a gene determines many aspects of its expression. RNA polymerase does not directly contact these gene-specific regulatory elements, however. Instead, proteins called transcription factors directly bind these sequences and interact with other components of the transcriptional machinery which in turn recruit RNA polymerase to the TSS. Accordingly, a typical transcription factor has two parts: a DNA-binding domain and an activation domain. DNA-binding domains are usually structured because they make stable contacts with specific regulatory sequences. In contrast, activations domains interact with the transcriptional machinery and are highly enriched in IDRs~\cite{Liu2006}. Decades of research have identified many of the proteins involved in these interactions as well as some common properties of their sequences~\cite{Ma1987, Sigler1988, Gerber1994, Arnold2018, Staller2018, Sanborn2021}. Despite this progress, however, the precise mechanisms by which transcription factors locate their DNA targets and recruit the transcriptional machinery to initiate expression of specific genes remains an unsolved problem in molecular biology. As disordered regions play a key role in this process, studies which broadly investigate the relationship between the sequence, structure, and function of IDRs therefore have significant implications for our understanding of gene regulation.
